---
title: "Cluster"
author: "Hüseyin Dalgıç"
date: "08 09 2020"
output: html_document
editor_options: 
  chunk_output_type: inline
---



#Libraries

```{r}
library(ggplot2)
library(tidyverse)
library(expss)
library(NbClust)
library(mclust)
library(gmodels)
library(funModeling)
library(cluster)    #kumeleme
library(factoextra) #gorsel
library(gridExtra)
library("fpc")
library(gridExtra)

```

# Explore data
```{r}

library(readr)
term_deposit_marketing_2020 <- read_csv("C:/Users/user/Desktop/R_Projeler/Hangikredi/term-deposit-marketing-2020.csv")


dataset<-term_deposit_marketing_2020

df<-dataset

summary(df)
str(df)


#replace "unknow" to NA
df[df == "unknown"] <- NA



#Converting all character variables to factor variables
df<-df %>%
  mutate_if(sapply(df, is.character), as.factor)



plot_num(df)
freq(df)



```

# Detecting interval range for numeric variables

## Age
```{r}
set.seed(1989)

age_class <- Mclust(df$age,verbose = FALSE)

summary(age_class)

df$age_range <- age_class$classification





table(df$age_range,df$age)

ggplot(df) +
  geom_density(aes(age, color=factor(age_range), fill=y))


df %>% 
    tab_cells(age) %>%
    tab_cols(age_range) %>% 
    tab_stat_fun("Toplam"=w_n, "Min"=w_min, "Max"=w_max, method=list) %>% 
    tab_pivot() %>% 
    set_caption("Table with summary statistics and significance marks.")


df<-df %>% 
  mutate(age_recode = case_when(
    age_range == 1  ~ "19-30",
    age_range >= 2 &age_range <= 5  ~ "31-40",
    age_range >= 6 &age_range <= 7  ~ "41-55",
    age_range >= 8  ~ "56+",
  )
)


```


## balance

```{r}


df<-df %>% 
  mutate(balance_recode = case_when(
    balance <= 0  ~ "(<=0)",
    balance > 0  ~ "(>0)"
  )
)



```


## day
```{r}
set.seed(1989)

day_class <- Mclust(df$day,verbose = FALSE)

summary(day_class)

df$day_range <- day_class$classification





table(df$day_range,df$day)

ggplot(df) +
  geom_density(aes(day, color=factor(day_range), fill=y))

ggplot(df) +
  geom_density(aes(day, fill=factor(day_range)))

df %>% 
    tab_cells(day) %>%
    tab_cols(day_range) %>% 
    tab_stat_fun("Toplam"=w_n, "Min"=w_min, "Max"=w_max, method=list) %>% 
    tab_pivot() %>% 
    set_caption("Table with summary statistics and significance marks.")


df<-df %>% 
  mutate(day_recode = case_when(
    day_range >= 1 &day_range <= 3  ~ "1-9",
    day_range == 4  ~ "10-14",
    day_range >= 5 &day_range <= 7  ~ "15-21",
    day_range == 8  ~ "22-26",
    day_range ==9  ~ "27-31"
  )
)



```




## duration
```{r}
set.seed(1989)

duration_class <- Mclust(df$duration,verbose = FALSE)

summary(duration_class)

df$duration_range <- duration_class$classification



ggplot(df) +
  geom_density(aes(duration, color=factor(duration_range), fill=y))

ggplot(df) +
  geom_density(aes(duration, fill=factor(duration_range)))

df %>% 
    tab_cells(duration) %>%
    tab_cols(duration_range) %>% 
    tab_stat_fun("Toplam"=w_n, "Min"=w_min, "Max"=w_max, method=list) %>% 
    tab_pivot() %>% 
    set_caption("Table with summary statistics and significance marks.")


df<-df %>% 
  mutate(duration_recode = case_when(
    duration_range ==1  ~ "0-28",
    duration_range >= 2 & duration_range <= 4  ~ "29-181",
    duration_range >= 5  ~ "257+",
  )
)



```




## campaign
```{r}

df<-df %>% 
  mutate(campaign_recode = case_when(
    campaign >=1 &  campaign <=2 ~ "1-2",
    campaign >2  ~ "2+"
  )
)



```


# Cluster

```{r}

summary(df)


df<-df %>% mutate_if(is.character, as.factor)

df_cluster<-df %>% select(duration,campaign)

df_cluster_copy<-df_cluster


```
## Kmenas

```{r}

k2 <- kmeans(scale(df_cluster), centers = 2, nstart = 25)


```


## Graphs


```{r}
fviz_cluster(k2, data = scale(df_cluster))


```
With Scatter chart
```{r}

df %>% as_tibble() %>%
  mutate(kumeler = k2$cluster, 
         codes = row.names(df)) %>%
  ggplot(aes(balance, age, color = factor(kumeler), label = codes)) +
  geom_text()




```

Graph of all cluster solutions
```{r}

k3 <- kmeans(scale(df_cluster) , centers = 3, nstart = 25)
k4 <- kmeans(scale(df_cluster) , centers = 4, nstart = 25)
k5 <- kmeans(scale(df_cluster) , centers = 5, nstart = 25)

p1 <- fviz_cluster(k2, geom = "point", data = scale(df_cluster)) + ggtitle("k=2")
p2 <- fviz_cluster(k3, geom = "point", data = scale(df_cluster)) + ggtitle("k=3")
p3 <- fviz_cluster(k4, geom = "point", data = scale(df_cluster)) + ggtitle("k=4")
p4 <- fviz_cluster(k5, geom = "point", data = scale(df_cluster)) + ggtitle("k=5")


grid.arrange(p1, p2, p3, p4, nrow = 2)


```


## Detecting cluster numbers with latent cluster analysis
```{r}
set.seed(1989)

lca_clust <- Mclust(scale(df_cluster),verbose = FALSE)
summary(lca_clust)

k7 <- kmeans(scale(df_cluster) , centers = 7, nstart = 25)


k7$centers


centers <- data.frame(duration=(data.frame(k7$centers)$duration*sd(df$duration) + mean(df$duration))/60,
                      campaign=data.frame(k7$centers)$campaign*sd(df$campaign) + mean(df$campaign))


fviz_cluster(k7, geom = "point", data = scale(df_cluster)) + ggtitle("k=7")




```

# Final cluster 
```{r}

k7 <- kmeans(scale(df_cluster) , centers = 3, nstart = 25)


k7$centers


centers <- data.frame(duration=(data.frame(k7$centers)$duration*sd(df$duration) + mean(df$duration))/60,
                      campaign=data.frame(k7$centers)$campaign*sd(df$campaign) + mean(df$campaign))


fviz_cluster(k7, geom = "point", data = scale(df_cluster)) + ggtitle("k=7")

df$clusters<-k7$cluster


df<-df %>% 
  mutate(clusters_recode = case_when(
    clusters ==1  ~ "Son gorusme 2,5 dakika , ortalama 15 kontakt sayisi",
    clusters ==2  ~ "Son gorusme 3 dakika, ortalama 2 kontakt sayisi",
    clusters ==3  ~ "Son gorusme 13 dakika, ortalama 2 kontakt sayisi ",
  )
)

df$clusters_recode<-factor(df$clusters_recode)

```


# Z Test

```{r}

lapply(select_if(df,is.factor), function(variable) {
  df %>%
    tab_cells(variable) %>%
    tab_cols(total(),y,clusters_recode) %>% 
    tab_stat_cpct(total_row_position = "above") %>% #yuzde esaplamasi gerektigini soyluyoruz
    tab_last_sig_cpct(min_base = 30) %>% #Anlamlilik testini uygulamasini min_base'i unutma default u 2 cunku 30 olmasi gerekir
    tab_pivot() %>% 
    set_caption("Table with summary statistics and significance marks.")
  })


```

# Executive Summary /Actions

For retired, student or management are more likely to accept offer. There 3 segments are splited by agetns's behaviors. We detected these segments with number of contacts performed during a campaign and last contact duration. If you look at test tables you can easily see segment 1:"Average last contact duration 13 min, average 2 contacts" has statiscial significant differences from other groups to subscribe to a term deposit. 



```{r}

cro_cpct(df$clusters_recode, df$y, total_label = "Sum", total_row_position = "above")


ggplot(df) +
  geom_bar(aes(x=y,y = (..count..)/sum(..count..)*100, fill=clusters_recode)) +
  ylab("Percent")



```

Demografig info 
Segment 1: blue-collar,retired,unemployed, /single,divorced/secondary 
Segment 2: technician,management  /married /tertiary  
Segment 3: retired,unemployed, /married,divorced/secondary 

```{r}



gp1<-ggplot() +
  geom_bar(data=df[df$clusters_recode=="Son gorusme 2,5 dakika , ortalama 15 kontakt sayisi",], mapping=aes(x=clusters_recode, y = (..count..)/sum(..count..)*100,  fill=job),position="dodge") +
    geom_bar(data=df[df$clusters_recode=="Son gorusme 3 dakika, ortalama 2 kontakt sayisi",], mapping=aes(x=clusters_recode, y = (..count..)/sum(..count..)*100,  fill=job),position="dodge") +
  geom_bar(data=df[df$clusters_recode=="Son gorusme 13 dakika, ortalama 2 kontakt sayisi ",], mapping=aes(x=clusters_recode, y = (..count..)/sum(..count..)*100,  fill=job),position="dodge")


gp2<-ggplot() +
  geom_bar(data=df[df$clusters_recode=="Son gorusme 2,5 dakika , ortalama 15 kontakt sayisi",], mapping=aes(x=clusters_recode, y = (..count..)/sum(..count..)*100,  fill=marital),position="dodge") +
    geom_bar(data=df[df$clusters_recode=="Son gorusme 3 dakika, ortalama 2 kontakt sayisi",], mapping=aes(x=clusters_recode, y = (..count..)/sum(..count..)*100,  fill=marital),position="dodge") +
  geom_bar(data=df[df$clusters_recode=="Son gorusme 13 dakika, ortalama 2 kontakt sayisi ",], mapping=aes(x=clusters_recode, y = (..count..)/sum(..count..)*100,  fill=marital),position="dodge")

gp3<-ggplot() +
  geom_bar(data=df[df$clusters_recode=="Son gorusme 2,5 dakika , ortalama 15 kontakt sayisi",], mapping=aes(x=clusters_recode, y = (..count..)/sum(..count..)*100,  fill=education),position="dodge") +
    geom_bar(data=df[df$clusters_recode=="Son gorusme 3 dakika, ortalama 2 kontakt sayisi",], mapping=aes(x=clusters_recode, y = (..count..)/sum(..count..)*100,  fill=education),position="dodge") +
  geom_bar(data=df[df$clusters_recode=="Son gorusme 13 dakika, ortalama 2 kontakt sayisi ",], mapping=aes(x=clusters_recode, y = (..count..)/sum(..count..)*100,  fill=education),position="dodge")




grid.arrange(gp1, gp2, gp3, nrow = 2)


```





Married people are more attended to be accept offers. Tertiary is alos powerfull info to say yes. Having a personal loan or house loan people are difficult to convince them to subscribe to a term deposit.

Februrary March April and October are goog time to call people. These monthes people are more likely to stay home and not to waste their money.

Whhen we examine day range, 10-14, 22-26 has significantly differences from other days. To keep duration at least 257 seconds has 2 times chance to subscribe to a term deposit.







